# Scenario VII: binomial distribution, Beta prior {#scenarioVII}

## Details

In this scenario, we revisit the case from [ScenarioVI](#variantVI), but are not assuming a point prior anymore.
We assume $p_C=0.3$ for the event rate in the control group. The parameter which can be varied is the event rate in the experimental group $p_E$ and we assume that $p_E \sim \text{Beta}(7,5)$.

In order to fulfill regulatory considerations, the type one error rate is still protected under the point prior $\delta=0$ at the level of significance $\alpha=0.025$.

The power constraint, however, needs to be modified. It is not senseful to compute the power as rejection probability under the full prior, because effect sizes less than a minimal clinically relevant effect do not show (sufficient) evidence against the null hypothesis. Therefore, we only consider possible event rates of at least $p_E=0.3$ and condition the prior on values $p_E>0.3$. In the following, the expected power should be at least $0.8$.


```{r}
# data distribution and priors
datadist   <- Binomial(rate_control=0.3, two_armed = TRUE)
H_0        <- PointMassPrior(.0, 1)
prior      <- ContinuousPrior(function(delta) dbeta(delta, shape1=7,shape2=5),
                              support = c(0,1),
                              tighten_support = TRUE)

# define constraints on type one error rate and expected power
alpha      <- 0.025
min_epower <- 0.8
toer_cnstr <- Power(datadist, H_0) <= alpha
epow_cnstr <- Power(datadist, condition(prior, c(0.3,1))) >= min_epower
```

## Variant II-1: Minimizing Expected Sample Size under Point Prior

### Objective

Expected sample size under the full prior is minimized, i.e., $\boldsymbol{E}\big[n(\mathscr{D})\big]$.

```{r}
ess <- ExpectedSampleSize(datadist, prior)
```

### Constraints

No additional constraints are considered.

### Initial Designs

For this example, the optimal one-stage, group-sequential, and generic two-stage designs are computed. The initial design for the one-stage case is determined heuristically and both the group sequential and the generic two-stage designs are optimized starting from the corresponding group-sequential design as computed by the `rpact` package.

```{r}
order <- 7L
# data frame of initial designs 
tbl_designs <- tibble(
    type    = c("one-stage", "group-sequential", "two-stage"),
    initial = list(
        OneStageDesign(200, 2.0),
        rpact_design(datadist, 0.3, 0.025, 0.8, TRUE, order),
        TwoStageDesign(rpact_design(datadist, 0.3, 0.025, 0.8, TRUE, order))) )
```
The order of integration is set to 7.

### Optimization

```{r}
tbl_designs <- tbl_designs %>% 
    mutate(
       optimal = purrr::map(initial, ~minimize(
          ess,
          subject_to(
              toer_cnstr,
              epow_cnstr
          ),
          
          initial_design = ., 
          opts           = opts)) )
```

### Test Cases

To avoid improper solutions, it is first verified that the maximum number of iterations was not exceeded in any of the three cases. 
