# Scenario IX: time-to-event endpoints

## Details
In this scenario, we visit a different scenario: Instead of continuous or binary endpoints, we now consider time-to-event endpoints. Therefore, we conduct a log-rank test, for which `adoptr` provides the normal approximation of the test statistics. In order to obtain a one-parametric statistics, we need to fix the event probability $d$ during the trial. We assume that a hazard ratio $\theta>1$ implies favorable results. The null-hypothesis is given by $\mathcal{H}_0:\theta\leq 1$. In both of the following two scenarios, we assume a constant event rate of $d=0.7$ and the type one error should be bounded by $\alpha<2.5$ and a power of $0.8$ is necessary. 

```{r}
datadist <- Survival(0.7,two_armed = TRUE)
H_0 <- PointMassPrior(1.0,1)
alpha <- 0.025
min_power <- 0.8
```

## Variant IX-1: Minimizing expected sample size under point prior

Under the alternative, we us a hazard ratio of $\theta=1.4$ for the point prior of the alternative. 

```{r}
prior <- PointMassPrior(1.4,1)

toer_cnstr <- Power(datadist,H_0)<=alpha
pow_cnstr <- Power(datadist, prior)>=min_power
```

### Objective
The expected sample size under the alternative is minimized.
```{r}
ess <- ExpectedSampleSize(datadist,prior)
```

### Constraints
No additional constraints besides type one and type two error are considered in this variant. 

### Initial Design
The initial design is chosen by using the `rpact` package.

```{r}

```


