# Scenario IX: Using further constraints

## General assumptions

`adoptr` offers the possibility to use a variety of further constraints. This chapter is dedicated to show some scenarios in which further conditions on the trial need to be met. We always assume a classical two-arm trial with normal test statistic and known variance. The null hypothesis is given by $\mathcal{H}_0: \delta \leq 0$ and the alternative is assumed to be an effect size of $\delta=0.5$ with point prior distribution. The type one error rate should be lower than $\alpha=0.025$ and the power at the alternative must be at least $0.9$. The objective function to be minimized is the expected sample size under the alternative, i.e. $\boldsymbol{E}\big[n(\mathcal{D})\big]$. 
```{r}
datadist <- Normal(two_armed=TRUE)
H_0 <- PointMassPrior(.0,1)
prior <- PointMassPrior(0.5,1)

alpha <- 0.025
min_power <- 0.9
toer_cnstr <- Power(datadist,H_0) <= alpha
pow_cnstr <- Power(datadist,prior) >= min_power

ess <- ExpectedSampleSize(datadist,prior)
```

## Constraint IX-1: Maximal Sample Size

### Details

Even though the expected sample size of adaptive two-stage designs is generally lower than the expected sample size of classical one-stage designs, it is possible that the sample size of two-stage designs exceeds the sample size of one-stage designs in some cases. Thus, it is not absurd to assume that the maximum sample size should have an upper bound. We say this upper bound is given by $1.4$ times the sample size of the one-stage design.

```{r}
init_onestage <- OneStageDesign(100,2)
opt_onestage <- minimize(ess,subject_to(toer_cnstr,pow_cnstr),initial_design = init_onestage)
print(evaluate(ess,opt_onestage$design))
```

Thus, the maximal sample size should not become larger than $118$.
```{r}
smplsize_cnstr <- MaximumSampleSize()<=118
```

### Initial Designs
As an initial design, we choose the `adoptr`-function `get_initial_data()`. 
```{r}
init_twostage <- get_initial_design(0.5,alpha,1-min_power,type=c("two-stage"),dist=datadist)
```

### Optimization
```{r}
opt_twostage_cnstr <- minimize(ess,subject_to(toer_cnstr,pow_cnstr,smplsize_cnstr),initial_design = init_twostage)
```

### Testcases

At first, we verify the type one error rate and power constraints by simulation:

```{r}
toer <- sim_pr_reject(opt_twostage_cnstr$design,0.0,datadist)
power <- sim_pr_reject(opt_twostage_cnstr$design,0.5,datadist)

testthat::expect_true(toer$prob<=alpha*(1+tol))
testthat::expect_true(power$prob>=min_power*(1-tol))
```

We now check that the maximal sample size is always smaller than $118$:
```{r}
testthat::expect_true(evaluate(MaximumSampleSize(),opt_twostage_cnstr$design)<=118)
```

Since we added a constraint to our minimization problem, the expected sample size of this design should be higher than the expected sample size without the constraint on the maximum sample size.

```{r}
opt_twostage <- minimize(ess,subject_to(toer_cnstr,pow_cnstr),initial_design = init_twostage)

testthat::expect_true(evaluate(ess,opt_twostage_cnstr$design)>= evaluate(ess,opt_twostage$design))
```



