# Scenario IX: time-to-event endpoints

## Details
In this scenario, we visit a different scenario: Instead of continuous or binary endpoints, we now consider time-to-event endpoints. Therefore, we conduct a log-rank test, for which `adoptr` provides the normal approximation of the test statistics. In order to obtain a one-parametric statistics, we need to fix the event probability $d$ during the trial. We assume that a hazard ratio $\theta>1$ implies favorable results. The null-hypothesis is given by $\mathcal{H}_0:\theta\leq 1$. In both of the following two scenarios, we assume a constant event rate of $d=0.7$ and the type one error should be bounded by $\alpha<2.5$ and a power of $0.8$ is necessary. 

```{r}
datadist <- Survival(0.7,two_armed = TRUE)
H_0 <- PointMassPrior(1.0,1)
alpha <- 0.025
min_power <- 0.8
```

## Variant IX-1: Minimizing expected sample size under point prior

Under the alternative, we us a hazard ratio of $\theta=1.4$ for the point prior of the alternative. 

```{r}
prior <- PointMassPrior(1.4,1)

toer_cnstr <- Power(datadist,H_0)<=alpha
pow_cnstr <- Power(datadist, prior)>=min_power
```

### Objective
The expected sample size under the alternative is minimized.
```{r}
ess <- ExpectedSampleSize(datadist,prior)
```

### Constraints
No additional constraints besides type one and type two error are considered in this variant. 

### Initial Design

The initial group-sequential and two-stage designs are chosen by using the `rpact` package, which calculates group-sequential designs. The one-stage design is chosen heuristically. 

```{r}
order <- 7L
# data frame of initial designs 
tbl_designs <- tibble(
    type    = c("one-stage", "group-sequential", "two-stage"),
    initial = list(
        OneStageDesign(150, 2.0),
        rpact_design(datadist, sig.level=0.025,power=0.8, two_armed=TRUE, order=order,hazardratio=1.4),
        TwoStageDesign(rpact_design(datadist, sig.level=0.025,power=0.8, two_armed=TRUE, 
                                    order=order,hazardratio=1.4))) )
```

The order of the integration is set to 7.

### Optimization

```{r}
tbl_designs <- tbl_designs %>% 
    mutate(
       optimal = purrr::map(initial, ~minimize(
         
          ess,
          subject_to(
              toer_cnstr,
              pow_cnstr
          ),
          
          initial_design = ., 
          opts           = opts)) )
```

### Test cases

We first check that the maximum number of iterations was not reached.

```{r}
tbl_designs %>% 
  transmute(
      type, 
      iterations = purrr::map_int(tbl_designs$optimal, 
                                  ~.$nloptr_return$iterations) ) %>%
  {print(.); .} %>% 
  {testthat::expect_true(all(.$iterations < opts$maxeval))}
```

Besides, we check whether the type one error was not exceeded and the necessary power is reached.

```{r}
tbl_designs %>% 
  transmute(
      type, 
      toer  = purrr::map(tbl_designs$optimal, 
                         ~sim_pr_reject(.[[1]], 1.0, datadist)$prob), 
      power = purrr::map(tbl_designs$optimal, 
                         ~sim_pr_reject(.[[1]], 1.4, datadist)$prob) ) %>% 
  unnest(., cols = c(toer, power)) %>% 
  {print(.); .} %>% {
  testthat::expect_true(all(.$toer  <= alpha * (1 + tol)))
  testthat::expect_true(all(.$power >= min_power * (1 - tol))) }
```

Since the degrees of freedom of the three design classes are ordered as
'two-stage' > 'group-sequential' > 'one-stage',
the expected sample sizes (under the alternative) should be ordered 
in reverse ('two-stage' smallest).
Additionally, expected sample sizes under both null and alternative
are computed both via `evaluate()` and simulation-based.

```{r}
ess0 <- ExpectedSampleSize(datadist, H_0)

tbl_designs %>% 
    mutate(
        ess      = map_dbl(optimal,
                           ~evaluate(ess, .$design) ),
        ess_sim  = map_dbl(optimal,
                           ~sim_n(.$design, 1.4, datadist)$n ),
        ess0     = map_dbl(optimal,
                           ~evaluate(ess0, .$design) ),
        ess0_sim = map_dbl(optimal,
                           ~sim_n(.$design, 1.0, datadist)$n ) ) %>% 
    {print(.); .} %>% {
    # sim/evaluate same under alternative?
    testthat::expect_equal(.$ess, .$ess_sim, 
                           tolerance = tol_n,
                           scale = 1)
    # sim/evaluate same under null?
    testthat::expect_equal(.$ess0, .$ess0_sim, 
                           tolerance = tol_n,
                           scale = 1)
    # monotonicity with respect to degrees of freedom
    testthat::expect_true(all(diff(.$ess) < 0)) }
```

The expected sample size under the alternative must be lower or equal than
the expected sample size of the initial `rpact` group-sequential design that
is based on the inverse normal combination test.

```{r}
testthat::expect_lte(
  evaluate(ess, 
             tbl_designs %>% 
                filter(type == "group-sequential") %>% 
                pull(optimal) %>% 
                .[[1]]  %>%
                .$design ),
    evaluate(ess, 
             tbl_designs %>% 
                filter(type == "group-sequential") %>% 
                pull(initial) %>% 
                .[[1]] ) )
```


